---
title: "DDSCASESTUDY1"
author: "Vanessa Nkongolo"
date: "2024-11-01"
output: html_document
---
```{r}
# Load libraries
library(ggplot2)
library(dplyr)
library(e1071)  # For Naive Bayes
library(caret)
library(class)

# Read the dataset
DF <- read.csv(choose.files())

# Select and process columns
CaseStudydata <- DF %>%
  select(Gender, MonthlyIncome, Distance = DistanceFromHome, NumCompWorked, 
         WorkLifeBalance, Rate = HourlyRate, Age, Department, JobSatisfaction, 
         EnvironmentSatisfaction, Overtime, Attrition) %>%
  mutate(
    JobSatisfactionOver3 = factor(ifelse(JobSatisfaction > 3, "Yes", "No")),
    WorkLifeBalanceOver3 = factor(ifelse(WorkLifeBalance > 3, "Yes", "No")),
    IsSalesDept = factor(ifelse(Department == "Sales", "Yes", "No")),
    DistanceFromHomeOver20 = factor(ifelse(Distance > 20, "Yes", "No")),
    Department = as.factor(Department),
    Overtime = as.factor(Overtime),
    Attrition = as.factor(Attrition)
  )
# Job Satisfaction by Attrition
ggplot(CaseStudydata, aes(x = Attrition, y = JobSatisfaction, fill = Attrition)) +
  geom_boxplot() +
  labs(title = "Job Satisfaction by Attrition", x = "Attrition Status", y = "Job Satisfaction Level") +
  scale_fill_manual(values = c("No" = "blue", "Yes" = "red")) +
  theme_minimal()

# Work-Life Balance by Attrition
ggplot(CaseStudydata, aes(x = factor(WorkLifeBalance), fill = Attrition)) +
  geom_bar(position = "fill") +
  scale_y_continuous(labels = scales::percent) +
  labs(title = "Attrition by Work-Life Balance", x = "Work-Life Balance Rating", y = "Percentage of Employees") +
  scale_fill_manual(values = c("No" = "blue", "Yes" = "red")) +
  theme_minimal()

# Distance from Home by Attrition
ggplot(CaseStudydata, aes(x = Attrition, y = Distance, fill = Attrition)) +
  geom_boxplot() +
  labs(title = "Attrition by Distance from Home", x = "Attrition Status", y = "Distance from Home (miles)") +
  scale_fill_manual(values = c("No" = "blue", "Yes" = "red")) +
  theme_minimal()
# Set a random seed for reproducibility
set.seed(42)

# Split data into training and testing sets
trainIndex <- createDataPartition(CaseStudydata$Attrition, p = 0.8, list = FALSE)
trainData <- CaseStudydata[trainIndex, ]
testData <- CaseStudydata[-trainIndex, ]
# Upsample the training data for balanced classes
trainData <- upSample(x = trainData[ , -which(names(trainData) == "Attrition")], y = trainData$Attrition)
colnames(trainData)[ncol(trainData)] <- "Attrition"

# Train Naive Bayes model
nb_model <- naiveBayes(Attrition ~ ., data = trainData)
print(nb_model)
# Upsample the training data for balanced classes
trainData <- upSample(x = trainData[ , -which(names(trainData) == "Attrition")], y = trainData$Attrition)
colnames(trainData)[ncol(trainData)] <- "Attrition"

# Train Naive Bayes model
nb_model <- naiveBayes(Attrition ~ ., data = trainData)
print(nb_model)
# Get predicted probabilities
prob_predictions <- predict(nb_model, testData, type = "raw")

# Adjust the threshold to optimize sensitivity and specificity
threshold <- 0.6
adjusted_predictions <- ifelse(prob_predictions[, "Yes"] > threshold, "Yes", "No")
adjusted_predictions <- as.factor(adjusted_predictions)

# Generate confusion matrix for adjusted predictions
confusion_matrix <- confusionMatrix(adjusted_predictions, testData$Attrition, positive = "Yes")
print(confusion_matrix)

# Extract accuracy, sensitivity, and specificity
accuracy <- confusion_matrix$overall['Accuracy']
sensitivity <- confusion_matrix$byClass['Sensitivity']
specificity <- confusion_matrix$byClass['Specificity']

cat("Accuracy:", accuracy, "\n")
cat("Sensitivity:", sensitivity, "\n")
cat("Specificity:", specificity, "\n")


# Case Study Prediction Code

```{r}
# Load necessary libraries
library(tidyverse)
library(dplyr)
library(e1071)  
library(caret)

# Load your original labeled data 
comp_set <- read.csv(choose.files())

# Verify that Department column exists before creating IsSalesDept
if("Department" %in% colnames(comp_set)) {
  comp_set <- comp_set %>%
    mutate(
      IsSalesDept = factor(ifelse(Department == "Sales", "Yes", "No")),
      WorkLifeBalanceOver3 = factor(ifelse(WorkLifeBalance > 3, "Yes", "No")),
      DistanceFromHomeOver20 = factor(ifelse(DistanceFromHome > 20, "Yes", "No")),
      Gender = as.factor(Gender),
      OverTime = as.factor(OverTime)
    )
} else {

  comp_set <- comp_set %>%
    mutate(
      WorkLifeBalanceOver3 = factor(ifelse(WorkLifeBalance > 3, "Yes", "No")),
      DistanceFromHomeOver20 = factor(ifelse(DistanceFromHome > 20, "Yes", "No")),
      Gender = as.factor(Gender),
      OverTime = as.factor(OverTime)
    )
}

# Select relevant columns and rename as needed
comp_set <- comp_set %>%
  select(ID, Gender, MonthlyIncome, DistanceFromHome, NumCompaniesWorked, 
         WorkLifeBalanceOver3, Rate = HourlyRate, Age, Department, 
         EnvironmentSatisfaction, OverTime, IsSalesDept, DistanceFromHomeOver20)

# Ensure factor levels match the training dataset
comp_set$Gender <- factor(comp_set$Gender, levels = levels(trainData$Gender))
comp_set$OverTime <- factor(comp_set$OverTime, levels = levels(trainData$OverTime))
comp_set$WorkLifeBalanceOver3 <- factor(comp_set$WorkLifeBalanceOver3, levels = levels(trainData$WorkLifeBalanceOver3))
if("IsSalesDept" %in% colnames(comp_set)) {
  comp_set$IsSalesDept <- factor(comp_set$IsSalesDept, levels = levels(trainData$IsSalesDept))
}
comp_set$DistanceFromHomeOver20 <- factor(comp_set$DistanceFromHomeOver20, levels = levels(trainData$DistanceFromHomeOver20))

# Use the Naive Bayes model to predict attrition
comp_set$Attrition <- predict(nb_model, comp_set)
comp_set$Attrition

# Filter Comp_Set
Filtered_comp_set <- comp_set %>% select(ID, Attrition)
Filtered_comp_set

# Define the file path within the CaseStudy1DDS directory
output_path <- "CaseStudy1CompSet_Predictions.csv"

# Save Filtered_comp_set as a CSV file
write.csv(Filtered_comp_set, file = output_path, row.names = FALSE)
